---
title: "Analyzing Car Data: Regression and Dummy Variables"
author: "Your Name"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

In this session, we’ll analyze a dataset of cars using linear regression. We’ll focus on creating dummy variables, interpreting models, and comparing group means.

---

## Step 1: Load the Data

We start by loading the dataset directly from GitHub and inspecting its structure.

```{r load-data}
# Load the dataset
kars <- read.table("https://raw.githubusercontent.com/dsoave/Datasets/refs/heads/main/mcars4.data")
dim(kars)  # Check dimensions
head(kars) # View first few rows
summary(kars) # Summary statistics
```

---

## Step 2: Data Transformation

Next, we modify the dataset by creating dummy variables and converting `Cntry` into a factor.

```{r transform-data}
n <- dim(kars)[1]
kars <- within(kars, {
  # Make dummy variables
  c1 <- numeric(n); c1[Cntry == 'Europ'] <- 1
  c2 <- numeric(n); c2[Cntry == 'Japan'] <- 1
  c3 <- numeric(n); c3[Cntry == 'US'] <- 1
  # Convert `Cntry` to a factor
  Cntry <- factor(Cntry)
}) # End of within kars

head(kars)
summary(kars)
```

---

## Step 3: Exploring Dummy Variables

We validate the dummy variables and calculate the mean fuel consumption for each country.

```{r validate-dummies}
# Validate dummy variables
with(kars, table(c1, Cntry))
with(kars, table(c2, Cntry))
with(kars, table(c3, Cntry))

# Take a look at mean fuel consumption for each country
with(kars, aggregate(lper100k, by = list(Cntry), FUN = mean))
# Must specify a LIST of grouping factors
```

---

## Step 4: Linear Models with Dummy Variables

We fit a linear regression model to test differences in fuel consumption by country.

```{r linear-models}
# Initial model using dummy variables
justcountry <- lm(lper100k ~ c1 + c2, data = kars)
summary(justcountry)

# Which means are different?
# Have t-tests. What about Europe vs. Japan?
# Test H0: beta1 = beta2
# A cheap way is to use a different reference category.
# R can make the dummy variables for you
is.factor(kars$Cntry)
# The factor Cntry has dummy vars built in. What are they?
contrasts(kars$Cntry) # Note alphabetical order

jc2 <- lm(lper100k ~ Cntry, data = kars)
summary(jc2)
```

---

## Step 5: Changing Reference Categories

We explore how changing reference categories affects the interpretation of results.

```{r reference-categories}
# Change reference category to 'US'
kars$Country <- kars$Cntry
contrasts(kars$Country) <- contr.treatment(3, base = 3)
summary(lm(lper100k ~ Country, data = kars))

# Names of dummy variables 1=Europe, 2=Japan could be nicer
colnames(contrasts(kars$Country)) <- c("Europe", "Japan")
contrasts(kars$Country)

summary(lm(lper100k ~ Country, data = kars))
```

---

## Step 6: Including Covariates

Finally, we include additional covariates like weight and length.

```{r covariates}
# Full model with covariates
fullmodel <- lm(lper100k ~ weight + length + Country, data = kars)
summary(fullmodel) # Look carefully at the signs!

# Reduced model for comparison
justsize <- lm(lper100k ~ weight + length, data = kars)
summary(justsize)

with(kars, cor(weight, length))

# I advise using anova ONLY to compare full and reduced models
anova(justsize, fullmodel) # Full vs reduced

# Test car size controlling for country too -- why not?
anova(justcountry, fullmodel)
```

---
